---
title: "Untitled"
author: "Keene Morrow"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)
```

## Grand Canyon GeoTIFF

```{r}
gc_dem <- raster(here("data", "gc_dem.tif"))

# look at geoTIFF using base R plot()
plot(gc_dem)

# check coordinate reference system (crs)
gc_dem@crs
# in UTM

# check extent (sometimes called bounding box)
gc_dem@extent
# uses meters

# get it into lat-long instead of m/UTM
# creating a WGS84 with lat-long metrics
wgs84 <- "+proj=longlat +zone=12 +datum=WGS84 +ellps=WGS84 +no_defs"

# reproject raster with new crs called 'wgs84' above
gc_reproj <- projectRaster(gc_dem, crs = wgs84, method = "bilinear")

#check extent
gc_reproj@extent
```

## Crop raster to smaller area
```{r}
# create new bounding box to crop reprojected geoTIFF
bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

# it needs a CRS still
# bounds@crs

# call the CRS of gc_dem
crs(bounds) <- crs(gc_reproj)

# bounds@crs

# Then crop gc_reproj by the new bounds polygon: 
gc_crop <- crop(gc_reproj, bounds)

# Look at it:
plot(gc_crop)

```

## Resample raster data using `aggregate()` function

```{r}
# aggregate cells by a factor of 10 (bigger factor, bigger resulting cells) , calculating mean
gc_agg <- raster::aggregate(gc_crop, fact = 10)

plot(gc_agg)
```

## How about in our ggplot security blanket?
```{r}
# make a data frame
# gc_df <- as.data.frame(gc_agg, xy = TRUE) %>%
#   rename(elev = layer)
# 
# ggplot(data = gc_df, aes(x = x, y = y)) +
#   geom_raster(aes(fill = elev)) +
#   coord_quickmap() +
#   theme_minimal() +
#   scale_fill_gradientn(colors = c(
#     "white",
#     "red",
#     "orange",
#     "yellow",
#     "green",
#     "dark green",
#     "blue",
#     "purple",
#     "black"
#     ))
```

### How to just select cells that match given criteria:

```{r}
# copy of dropped dem to mess with
gc_hab <- gc_crop

# set any cells outside of [1000 - 1500] to NA 
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA

# plot
plot(gc_hab)

```

### Does it work with the interactive mapping?
(That's half the point)
```{r}
tmap_mode("view")

tm_shape(gc_hab) +
  tm_raster(legend.show = FALSE, palette = "plasma")
```

### Kriging rain in Kansas
For a particularly rainy day in 2017

Read in Kansas counties shapefile data
```{r}
ks_counties <- read_sf(here("data",
                            "ks_counties",
                            "ks_counties_shapefile.shp"))

# will try to plot all attributes
plot(ks_counties)

# we're back in sf, so we're now using sf functions
# here, check the CRS
st_crs(ks_counties)

# there isn't one, so let's set the CRS to EPGS 4326
st_crs(ks_counties) <- 4326

# it now has a CRS
st_crs(ks_counties)

# no longers looks stretched
plot(ks_counties)
```

## Now in ggplot
```{r}
ggplot(data = ks_counties) +
  geom_sf()
```

Now to read in the rainfall data
```{r}
ks_rain <- read_csv(here("data", "ks_rain.csv")) %>%
  clean_names()
```

but R doesn't know that this is spatial data
time to fix that

```{r}
ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326)

ggplot() +
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf,
          aes(color = amt, size = amt),
          show.legend = FALSE)

```

# Kriging to Predict Rainfall

We'll do this using gstat, which doesn't work well with sf objects

```{r}
ks_sp <- as_Spatial(ks_sf)

class(ks_sp)
```

### Make a spatial pixels grid to make predictions over
```{r}
# determine the extent
bbox(ks_sp)

# set up the grid components
lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6,-102, length.out = 200)

# now to build the grid in three steps
grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon", "lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

plot(grid_sp)
```

#### Make a variogram from the rainfall data: 
```{r}
# Create the variogram:
ks_vgm <- variogram(amt ~ 1, data = ks_sp)

# Look at it: 
plot(ks_vgm)

# Estimates:
# nugget = 0.2
# sill = 0.8
# range = 200

# Spherical model looks good
ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2,
                                                psill = 0.8,
                                                range = 200,
                                                model = "Sph"))
plot(ks_vgm, ks_vgm_fit)

# Exponential model looks good
ks_vgm_fit_exp <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2,
                                                psill = 0.8,
                                                range = 200,
                                                model = "Exp"))
plot(ks_vgm, ks_vgm_fit_exp)

# Gaussian model looks concerning
ks_vgm_fit_gau <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2,
                                                psill = 0.8,
                                                range = 200,
                                                model = "Gau"))
plot(ks_vgm, ks_vgm_fit_gau)
```

Call `ks_vgm_fit` in the console to see:
        model     psill    range
      1   Nug 0.1021677   0.0000
      2   Sph 0.9537363 235.1416

### Kriging
```{r}
ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)
```

We can look at this with spplot
```{r}
spplot(ks_krige, "var1.pred")
```

### Make a data frame of the kriged predictions 
```{r}
ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(long = coords.x1,
         lat = coords.x2)
```

Converting this df to sf object
```{r}
rain_sf <- st_as_sf(ks_df, coords = c("long", "lat"), crs = 4326)

ggplot(rain_sf) +
  geom_sf(aes(color = var1.pred))
```

### Crop to the actual KS outline

```{r}
# get Kansas from all states and transform to CRS of existing data

ks <- read_sf(dsn = here("data", "states"),
              layer = "cb_2017_us_state_20m") %>%
  clean_names() %>%
  dplyr::select(name) %>%
  dplyr::filter(name == "Kansas") %>%
  st_transform(crs = 4326)

plot(ks)
```

```{r}
# find the intersection of the two:
rain_sf_ks <- st_intersection(rain_sf, ks)

plot(rain_sf_ks)
```

And plot in ggpiot
```{r}
ggplot(data = rain_sf_ks) +
  geom_sf(aes(color = var1.pred))
```

